Party tricks for PostgreSQL \\ perf, ftrace and bpftrace

### 2022-10-21 pgconfeu2022/presentation.tex
Editing

\usepackage[utf8]{inputenc}
\usepackage[russian, english]{babel}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{xcolor}
\usepackage[document]{ragged2e}
\usepackage[export]{adjustbox}
\usepackage{fontawesome5}
\usepackage{enumitem}
\usepackage{minted}
\usemintedstyle{tango}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usetikzlibrary{mindmap}
\usetikzlibrary{shapes.misc, fit}
\usetikzlibrary{spy, decorations, decorations.pathmorphing, decorations.pathreplacing, backgrounds}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{verbatim}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\usetikzlibrary{trees}
\usepackage{datetime}
\usepackage{hyperref}
\usepackage{forloop}
\usepackage{csquotes}
\usetikzlibrary{tikzmark}
\usetikzlibrary{arrows.meta}

@@

@@

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{c}
*dovacuum =
	force_vacuum ||
	(vactuples > vacthresh) ||
	(vac_ins_base_thresh >= 0 &&
		instuples > vacinsthresh);
        \end{minted}
    \end{center}
\end{frame}
\note{
	Filter doesn't work with comparisons, requires post-processing,
	complicated. For such cases it's good to have a stable usdt.
}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[russian, english]{babel}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{listings}
\usepackage{marvosym}
\usepackage{color}
\usepackage{xcolor}
\usepackage[document]{ragged2e}
\usepackage[export]{adjustbox}
\usepackage{fontawesome5}
\usepackage{enumitem}
\usepackage{minted}
\usemintedstyle{tango}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usetikzlibrary{mindmap}
\usetikzlibrary{shapes.misc, fit}
\usetikzlibrary{spy, decorations, decorations.pathmorphing, decorations.pathreplacing, backgrounds, decorations.text}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{verbatim}
\usepackage{smartdiagram}
\usesmartdiagramlibrary{additions}
\usetikzlibrary{trees}
\usepackage{datetime}
\usepackage{hyperref}
\usepackage{forloop}
\usepackage{csquotes}
\usetikzlibrary{tikzmark}
\usetikzlibrary{arrows.meta}

@@

\section{Limitations}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large, escapeinside=::]{c}
*dovacuum =
	force_vacuum ||
	(vactuples > vacthresh) ||
	(vac_ins_base_thresh >= 0 &&
		:\colorbox{red!20}{instuples > vacinsthresh)};:
        \end{minted}
    \end{center}
\end{frame}
\note{
	Filter doesn't work with comparisons, requires post-processing,
	complicated. For such cases it's good to have a stable usdt.
}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{itemize}[label={}]
            \item<+-> {Context switch overhead}
            \item<+-> {Debugging symbols}
            \item<+-> {Incorporate into infrastructure?}
            \item<+-> {Privileges?}
        \end{itemize}
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{c}
BPF_OBJ_PIN
    Pin an eBPF program or map referred
    by the specified *bpf_fd* to the
    provided *pathname* on the filesystem.
        \end{minted}
    \end{center}
\end{frame}
\note{
    BPFD
}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{shell-session}
$ cat events/synthetic/query_latency/hist
# event histogram

hitcount: 1
{ lat_us: 223,
  query_string: select a from test1;}

hitcount: 1
{ lat_us: 314,
  query_string: select b from test2;}
        \end{minted}
    \end{center}
\end{frame}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{shell-session}
$ perf probe -x bin/postgres \
    exec_simple_query query_string:string

$ perf probe -x bin/postgres \
    exec_simple_query%return
        \end{minted}
    \end{center}
\end{frame}

@@

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
    \begin{overprint}[14.5cm]
        \onslide<1>
		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
synthetic_events
query_latency u64 lat_us; char[] query_string
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/probe_postgres/exec_simple_query/trigger
hist:keys=common_pid:t0=common_timestamp.usecs
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/probe_postgres/exec_simple_query__return/trigger
hist:keys=common_pid:lat_us=common_timestamp.usecs-$t0:
   onmatch(probe_postgres.exec_simple_query).
   trace(query_latency,$lat_us,query_string_string)
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/synthetic/query_latency/trigger
hist:keys=lat_us,query_string:sort=lat_us
        \end{minted}

        \onslide<2>
        \begin{minted}[fontsize=\large, bgcolor=red!20]{shell-session}
synthetic_events
query_latency u64 lat_us; char[] query_string
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/probe_postgres/exec_simple_query/trigger
hist:keys=common_pid:t0=common_timestamp.usecs
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/probe_postgres/exec_simple_query__return/trigger
hist:keys=common_pid:lat_us=common_timestamp.usecs-$t0:
   onmatch(probe_postgres.exec_simple_query).
   trace(query_latency,$lat_us,query_string_string)
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/synthetic/query_latency/trigger
hist:keys=lat_us,query_string:sort=lat_us
        \end{minted}

        \onslide<3>
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
synthetic_events
query_latency u64 lat_us; char[] query_string
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=red!20]{shell-session}
events/probe_postgres/exec_simple_query/trigger
hist:keys=common_pid:t0=common_timestamp.usecs
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/probe_postgres/exec_simple_query__return/trigger
hist:keys=common_pid:lat_us=common_timestamp.usecs-$t0:
   onmatch(probe_postgres.exec_simple_query).
   trace(query_latency,$lat_us,query_string_string)
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/synthetic/query_latency/trigger
hist:keys=lat_us,query_string:sort=lat_us
        \end{minted}

        \onslide<4>
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
synthetic_events
query_latency u64 lat_us; char[] query_string
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/probe_postgres/exec_simple_query/trigger
hist:keys=common_pid:t0=common_timestamp.usecs
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=red!20]{shell-session}
events/probe_postgres/exec_simple_query__return/trigger
hist:keys=common_pid:lat_us=common_timestamp.usecs-$t0:
   onmatch(probe_postgres.exec_simple_query).
   trace(query_latency,$lat_us,query_string_string)
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/synthetic/query_latency/trigger
hist:keys=lat_us,query_string:sort=lat_us
        \end{minted}

        \onslide<5>
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
synthetic_events
query_latency u64 lat_us; char[] query_string
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/probe_postgres/exec_simple_query/trigger
hist:keys=common_pid:t0=common_timestamp.usecs
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=white]{shell-session}
events/probe_postgres/exec_simple_query__return/trigger
hist:keys=common_pid:lat_us=common_timestamp.usecs-$t0:
   onmatch(probe_postgres.exec_simple_query).
   trace(query_latency,$lat_us,query_string_string)
        \end{minted}

		\vspace{-0.73cm}
        \begin{minted}[fontsize=\large, bgcolor=red!20]{shell-session}
events/synthetic/query_latency/trigger
hist:keys=lat_us,query_string:sort=lat_us
        \end{minted}

    \end{overprint}

    \end{center}
\end{frame}

\section{Development}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{shell-session}
$ perf record -a -g --event=mem:0x7ff3bc1f9000/8:rw
        \end{minted}
    \end{center}
\end{frame}
\note{share memory from postmaster maps}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{text}
$ bpftrace -e 'kprobe:__x64_sys_pwrite64
    /comm == "postgres"/ { override(-12) }' --unsafe
Attaching 1 probe...
        \end{minted}

        \begin{minted}[fontsize=\large]{text}
=# create table test();
PANIC:  XX000: could not write to
log file 00000001000000000000002B at offset 11444224,
length 106496: Cannot allocate memory
LOCATION:  XLogWrite, xlog.c:2233
        \end{minted}
    \end{center}
\end{frame}
\note{ENOMEM, only for those functions that are markes safe to fail.}

\section{Tips \& tricks}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{shell-session}
$ perf report --header

# ========
# captured on    : Wed Oct 12 17:07:32 2022
# header version : 1
# data offset    : 504
# data size      : 1392968
# feat offset    : 1393472
# arch : x86_64
# cpudesc : Intel(R) Core(TM) i7-10610U CPU @ 1.80GHz
# ...
        \end{minted}
    \end{center}
\end{frame}
\note{ENOMEM, only for those functions that are markes safe to fail.}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{shell-session}
$ perf archive

  Now please run:
  tar xvf perf.data.tar.bz2 -C ~/.debug
        \end{minted}
    \end{center}
\end{frame}
\note{Archives all available symbols}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{shell-session}
$ perf record --overwrite --switch-output=signal
$ kill -SIGUSR2 <perf-pid>
        \end{minted}
    \end{center}
\end{frame}
\note{Daemonized perf, optionally specify ring buffer size with -m 1M to
    control memory better}

\begin{frame}[fragile]{}
    \frametitle{}
    \begin{center}
        \begin{minted}[fontsize=\large]{shell-session}
$ perf record -p <pid> -e intel_pt//u --filter\
 'filter tts_buffer_heap_materialize @ bin/postgres'
        \end{minted}
    \end{center}
\end{frame}
\note{Duplicated functions?}

\begin{frame}
    \frametitle{}
    \begin{center}
		\begin{tikzpicture}[
		  decoration={
			reverse path,
			text effects along path,
			text={perf record -e probe\_postgres:GetNewTransactionId perf script bpftrace -l 'usdt:bin/postgres:*' bpftrace -p 31715 -f json -e 'usdt:bin/postgres:postgresql:checkpoint\_\_start printf("Checkpoint, Flags d", arg0);' perf probe -x bin/postgres -L GetNewTransactionId perf probe -x bin/postgres -V GetNewTransactionId:187 perf probe -x bin/postgres GetNewTransactionId:187 nxids perf record -e probe\_postgres:GetNewTransactionId\_L187\_2 perf script cat events/synthetic/query\_latency/hist perf probe -x bin/postgres exec\_simple\_query query\_string:string perf probe -x bin/postgres exec\_simple\_query return echo 'query\_latency u64 lat\_us; char[] query\_string' > synthetic\_events echo 'hist:keys=common\_pid:ts0=common\_timestamp.usecs' > events/probe\_postgres/exec\_simple\_query/trigger echo 'hist:keys=common\_pid:lat\_us=common\_timestamp.usecs-ts0:onmatch(probe\_postgres.exec\_simple\_query).trace(query\_latency,lat\_us,query\_string\_string)' > events/probe\_postgres/exec\_simple\_query\_\_return/trigger echo 'hist:keys=lat\_us,query\_string:sort=lat\_us' > events/synthetic/query\_latency/trigger perf record -a -g --event=mem:0x7ff3bc1f9000/8:rw bpftrace -e 'kprobe:\_\_x64\_sys\_pwrite64 comm == "postgres"  override(-12) ' --unsafe perf report --header perf archive perf record --overwrite --switch-output=signal kill -SIGUSR2 <perf-pid> perf record -p <pid> -e intel\_pt//u --filter 'filter tts\_buffer\_heap\_materialize @ bin/postgres'},
			text effects/.cd,
			  text along path,
			  character count=\i, character total=\n,
			  characters={scale=1-\i/\n}
			}
		]
		\draw [decorate] (0,0)
			\foreach \i [evaluate={\r=(\i/2000)^2;}] in {0,5,...,4880}{ -- (\i:\r)};
		\end{tikzpicture}
    \end{center}
\end{frame}
